Grade Retention and Mixed Race

### 2022-04-04 products/paper/main.qmd
Replace figure of sample sizes with descriptive stats table

the cross-tabulation of parent's races leads to ten distinct biracial categories that respondents may fall into. The total sample size for each of these biracial categories is shown in @fig-sample-size. The sample sizes are quite large for many of the groups, suggesting substantial statistical power in determining the relative outcomes for these groups. Even among the smallest group of Indigenous/Asian, the sample size is substantially greater than one would see in typical survey data ($n=519$).

```{r}
#| label: fig-sample-size
#| fig-cap: Sample size of biracial respondents combined across American Community Survey data.
acs %>%
  filter(str_detect(race, "/")) %>%
  ggplot(aes(x=reorder(race, race, length)))+
  geom_bar()+
  coord_flip()+
  labs(x=NULL, y="number of respondents")+
  scale_y_continuous(labels = scales::comma)+
  theme_bw()
```

@@

To account for this grade-specific decline in the models I include an interaction between a linear year term and current grade. Sensitivity analysis showed that this functional form was preferred by BIC to models with no interaction and a model with interaction terms that treated year as a categorical variable. (SHOULD I BE CONSIDERING THE SAME FOR STATE?)

>>>>

the cross-tabulation of parent's races leads to ten distinct biracial categories that respondents may fall into. The total sample size for each of these biracial categories is shown in @tbl-descriptive, along with several descriptive statistics. The sample sizes are quite large for many of the groups, suggesting substantial statistical power in determining the relative outcomes for these groups. Even among the smallest group of Indigenous/Asian, the sample size is substantially greater than one would see in typical survey data.

```{r}
#| label: tbl-descriptive
#| tbl-cap: Descriptive statistics by race. Shading indicates biracial group. Groups ordered by sample size.
tbl_summary <- acs %>%
  group_by(race) %>%
  summarize(n=n(), 
            held_back=percent(mean(below_exp_grade), accuracy=0.01),
            mean_age=round(mean(age), 1),
            f_born=percent(mean(foreign_born),  accuracy=0.1),
            par_fb=percent(mean(foreign_born_father | 
                                  foreign_born_mother), accuracy=0.1),
            fam_inc=dollar(round(mean(family_income), 0)),
            own_home=percent(mean(own_home), accuracy=0.1),
            par_ba=percent(mean(degree_father=="BA" | 
                                  degree_father=="G" | 
                                  degree_mother=="BA" | 
                                  degree_mother=="G"), accuracy=0.1)
  ) %>%
  arrange(desc(n)) %>%
  mutate(n=comma(n))

index_biracial <- which(str_detect(tbl_summary$race, "/"))

tbl_summary %>%
  kbl(booktabs = TRUE,
      linesep = NULL,
      align = c("l",rep("r", ncol(tbl_summary)-1)),
      col.names=c("race",
                  "n",
                  "clearly below expected grade", 
                  "mean age",
                  "foreign-born",
                  "either parent foreign-born",
                  "mean family income",
                  "own home",
                  "either parent, college degree"
      )) %>%
  column_spec(3:10, width = "1.6cm") %>%
  row_spec(index_biracial, background = "#D3D3D3") %>%
  landscape()
```

```{r}
#| eval: false
#| label: fig-sample-size
#| fig-cap: Sample size of biracial respondents combined across American Community Survey data.
acs %>%
  filter(str_detect(race, "/")) %>%
  ggplot(aes(x=reorder(race, race, length)))+
  geom_bar()+
  coord_flip()+
  labs(x=NULL, y="number of respondents")+
  scale_y_continuous(labels = scales::comma)+
  theme_bw()
```

@@

To account for this grade-specific decline in the models I include an interaction between a linear year term and current grade. Sensitivity analysis showed that this functional form was preferred by BIC to models with no interaction and a model with interaction terms that treated year as a categorical variable. (==SHOULD I BE CONSIDERING THE SAME FOR STATE?==)
